STAT 253 Final Project - Classification Models

### 2022-11-02 Final_Project_Classification_Models.Rmd
Tentative Forest Model

```{r}
library(readr)
imdb_top_1000 <- read_csv("~/Desktop/Statistical Machine Learning/R Files/Final Project/imdb_top_1000.csv")
```

@@

```{r}
imdb_clean <- imdb_top_1000 %>%
  select(-Poster_Link, -Certificate) %>%
  cSplit("Genre", sep = ",", direction = "wide") %>%
  mutate(Gross = log(Gross))

@@

imdb_clean$Runtime <- runtime_clean
```

# Logistic Regression

```{r}
# Logistic Regression Model Spec
logistic_spec <- logistic_reg() %>%
    set_engine('glm') %>%
    set_mode('classification')

# Recipe
logistic_rec <- recipe(___ ~ ., data = imdb_clean) %>%
    step_normalize(all_numeric_predictors()) %>% 
    step_dummy(all_nominal_predictors())

# Workflow (Recipe + Model)
log_wf <- workflow() %>% 
    add_recipe(logistic_rec) %>%
    add_model(logistic_spec)

# Fit Model to Training Data
log_fit <- fit(log_wf, data = imdb_clean)
```

### Examining the Logistic Regression Model

```{r}
# Print out Coefficients
log_fit %>% tidy()

# Get Exponentiated coefficients + CI
log_fit %>% tidy() %>%
  mutate(OR.conf.low = exp(estimate - 1.96*std.error), OR.conf.high = exp(estimate + 1.96*std.error)) %>% # do this first
  mutate(OR = exp(estimate))
```

### Making Prediction on Logistic Model

```{r}
# Make soft (probability) predictions
predict(log_fit, new_data = ___, type = "prob")

# Make hard (class) predictions (using a default 0.5 probability threshold)
predict(log_fit, new_data = ___, type = "class")
```

### Evaluating Logistic Model on Training Data

```{r}
# Soft predictions
logistic_output <-  data %>%
  bind_cols(predict(log_fit, new_data = data, type = 'prob'))

# Hard predictions (you pick threshold)
logistic_output <- logistic_output %>%
  mutate(.pred_class = make_two_class_pred(.pred_REFLEVEL, levels(outcome), threshold = .8)) #Try changing threshold (.5, 0, 1, .2, .8)

# Visualize Soft Predictions
logistic_output %>%
  ggplot(aes(x = outcome, y = .pred_OUTCOME)) +
  geom_boxplot() + 
  geom_hline(yintercept = 0.5, color='red') +  # try changing threshold
  labs(y = 'Predicted Probability of Outcome', x = 'Observed Outcome') +
  theme_classic()
```

# Calculating Evaluation Metrics

```{r}
# Confusion Matrix
logistic_output %>%
  conf_mat(truth = outcome, estimate = .pred_class)

log_metrics <- metric_set(sens, yardstick::spec, accuracy) # these metrics are based on hard predictions

#sens: sensitivity = chance of correctly predicting second level, given second level (Yes)
#spec: specificity = chance of correctly predicting first level, given first level (No)
#accuracy: accuracy = chance of correctly predicting outcome

logistic_output %>% 
  log_metrics(estimate = .pred_class, truth = outcome, event_level = "second") # set second level of outcome as "success"
```

### ROC Curve

```{r}
logistic_roc <- logistic_output %>% 
    roc_curve(outcome, .pred_OUTCOME, event_level = "second") # set second level of outcome as "success"

autoplot(logistic_roc) + theme_classic()
```

### Evaluation Metrics Using CV

```{r}
# CV Fit Model
log_cv_fit <- fit_resamples(
    log_wf, 
    resamples = data_cv10,
    metrics = metric_set(sens, yardstick::spec, accuracy, roc_auc),
    control = control_resamples(save_pred = TRUE, event_level = 'second'))  # you need predictions for ROC calculations

collect_metrics(log_cv_fit) #default threshold is 0.5
```

# Logistic Model with LASSO

```{r}
set.seed(123)

# Make sure you set reference level (to the outcome you are NOT interested in)
spam <- spam %>%
  mutate(spam = relevel(factor(spam), ref='not_spam')) #set reference level

data_cv10 <- vfold_cv(spam, v = 10)

# Logistic LASSO Regression Model Spec
logistic_lasso_spec_tune <- logistic_reg() %>%
    set_engine('glmnet') %>%
    set_args(mixture = 1, penalty = tune()) %>%
    set_mode('classification')

# Recipe
logistic_rec <- recipe(_____, data = spam) %>%
    step_normalize(all_numeric_predictors()) %>% 
    step_dummy(all_nominal_predictors())

# Workflow (Recipe + Model)
log_lasso_wf <- workflow() %>% 
    add_recipe(logistic_rec) %>%
    add_model(logistic_lasso_spec_tune)

# Tune Model (trying a variety of values of Lambda penalty)
penalty_grid <- grid_regular(
  penalty(range = c(__,__)), #log10 transformed 
  levels = 100)

tune_output <- tune_grid( 
  log_lasso_wf, # workflow
  resamples = data_cv10, # cv folds
  metrics = metric_set(roc_auc,accuracy),
  control = control_resamples(save_pred = TRUE, event_level = 'second'),
  grid = penalty_grid # penalty grid defined above
)

# Visualize Model Evaluation Metrics from Tuning
autoplot(tune_output) + theme_classic()
```

### Inspecting the Model

```{r}
# Visualize Model Evaluation Metrics from Tuning
autoplot(tune_output) + theme_classic()
```

```{r}
# Select Penalty
best_se_penalty <- select_by_one_std_err(tune_output, metric = 'roc_auc', desc(penalty)) # choose penalty value based on the largest penalty within 1 se of the highest CV roc_auc
best_se_penalty

# Fit Final Model
final_fit_se <- finalize_workflow(log_lasso_wf, best_se_penalty) %>% # incorporates penalty value to workflow 
    fit(data = spam)

final_fit_se %>% tidy()
```

```{r}
# Comment on Variable Importance

glmnet_output <- final_fit_se %>% extract_fit_engine()

# Create a boolean matrix (predictors x lambdas) of variable exclusion
bool_predictor_exclude <- glmnet_output$beta==0

# Loop over each variable
var_imp <- sapply(seq_len(nrow(bool_predictor_exclude)), function(row) {
    this_coeff_path <- bool_predictor_exclude[row,]
    if(sum(this_coeff_path) == ncol(bool_predictor_exclude)){ return(0)}else{
    return(ncol(bool_predictor_exclude) - which.min(this_coeff_path) + 1)}
})

# Create a dataset of this information and sort
var_imp_data <- tibble(
    var_name = rownames(bool_predictor_exclude),
    var_imp = var_imp
)
var_imp_data %>% arrange(desc(var_imp))
```

# Evaluation Metrics

```{r}

# CV results for "best lambda"
tune_output %>%
    collect_metrics() %>%
    filter(penalty == best_se_penalty %>% pull(penalty))

# Count up number of spam and not_spam emails in the training data
spam %>%
    count(spam) # Name of the outcome variable goes inside count()

# Compute the NIR
```

### Final Model

```{r}
# Soft Predictions on Training Data
final_output <- final_fit_se %>% predict(new_data = spam, type='prob') %>% bind_cols(spam)
```

```{r}
# Use soft predictions
final_output %>%
    roc_curve(spam,.pred_spam,event_level = 'second') %>%
    autoplot()
```

```{r}
# thresholds in terms of reference level
threshold_output <- final_output %>%
    threshold_perf(truth = spam, estimate = .pred_not_spam, thresholds = seq(0,1,by=.01))

# J-index v. threshold for not_spam
threshold_output %>%
    filter(.metric == 'j_index') %>%
    ggplot(aes(x = .threshold, y = .estimate)) +
    geom_line() +
    labs(y = 'J-index', x = 'threshold') +
    theme_classic()

threshold_output %>%
    filter(.metric == 'j_index') %>%
    arrange(desc(.estimate))

# Distance v. threshold for not_spam

threshold_output %>%
    filter(.metric == 'distance') %>%
    ggplot(aes(x = .threshold, y = .estimate)) +
    geom_line() +
    labs(y = 'Distance', x = 'threshold') +
    theme_classic()

threshold_output %>%
    filter(.metric == 'distance') %>%
    arrange(.estimate)

log_metrics <- metric_set(accuracy,sens,yardstick::spec)

final_output %>%
    mutate(.pred_class = make_two_class_pred(.pred_not_spam, levels(spam), threshold = ___ )) %>%
    log_metrics(truth = spam, estimate = .pred_class, event_level = 'second')
```

@@

```{r}
# Model Specification
rf_spec <- rand_forest() %>%
  set_engine(engine = 'ranger') %>% 
  set_args(mtry = NULL, # size of random subset of variables; default is floor(sqrt(number of total predictors))
           trees = 1000, # Number of trees
           min_n = 2,
           probability = FALSE, # FALSE: get hard predictions (not needed for regression)
           importance = 'impurity') %>% # we'll come back to this at the end
  set_mode('classification') # change this for regression

# Recipe
data_rec <- recipe(class ~ ., data = land)

# Workflows
data_wf_mtry2 <- workflow() %>%
  add_model(rf_spec %>% set_args(mtry = 2)) %>%
  add_recipe(data_rec)

## Create workflows for mtry = 12, 74, and 147

data_wf_mtry12 <-

data_wf_mtry74 <-

data_wf_mtry147 <- 
```

```{r}
# Fit Models
set.seed(123) # make sure to run this before each fit so that you have the same 1000 trees
data_fit_mtry2 <- fit(data_wf_mtry2, data = land)

# Fit models for 12, 74, 147
set.seed(123) 
data_fit_mtry12 <-

set.seed(123)
data_fit_mtry74 <-

set.seed(123) 
data_fit_mtry147 <- 
```

```{r}
# Custom Function to get OOB predictions, true observed outcomes and add a user-provided model label
rf_OOB_output <- function(fit_model, model_label, truth){
    tibble(
          .pred_class = fit_model %>% extract_fit_engine() %>% pluck('predictions'), #OOB predictions
          class = truth,
          label = model_label
      )
}

#check out the function output
rf_OOB_output(data_fit_mtry2,2, land %>% pull(class))
```

@@

data_rf_OOB_output <- bind_rows(
    rf_OOB_output(data_fit_mtry2,2, land %>% pull(class)),
    rf_OOB_output(data_fit_mtry12,12, land %>% pull(class)),
    rf_OOB_output(data_fit_mtry74,74, land %>% pull(class)),
    rf_OOB_output(data_fit_mtry147,147, land %>% pull(class))
)

@@

```{r}
rf_OOB_output(data_fit_mtry12,12, land %>% pull(class)) %>%
    conf_mat(truth = class, estimate= .pred_class)
```

@@

model_output <-data_fit_mtry12 %>% 
    extract_fit_engine()

model_output %>% 
    vip(num_features = 30) + theme_classic() #based on impurity

@@

model_output2 <- data_wf_mtry12 %>% 
  update_model(rf_spec %>% set_args(importance = "permutation")) %>% #based on permutation
  fit(data = land) %>% 
    extract_fit_engine()

model_output2 %>% 
    vip(num_features = 30) + theme_classic()

>>>>

```{r}
imdb_top_1000 <- read_csv("~/Desktop/Statistical Machine Learning/R Files/Final Project/imdb_top_1000_updated.csv")
```

@@

```{r}
imdb_clean <- imdb_top_1000 %>%
  cSplit("Genre", sep = ",", direction = "wide") %>%
  mutate(Gross = log(revenue))

@@

imdb_clean$Runtime <- runtime_clean

imdb_clean_class <- imdb_clean %>%
  mutate(success_ratio = revenue/budget) %>%
  mutate(flop = as.factor(ifelse(success_ratio > 1, 'FALSE', 'TRUE'))) %>%
  filter(!is.na(flop))

```

@@

```{r}
# Model Specification
rf_spec <- rand_forest() %>%
  set_engine(engine = 'ranger') %>% 
  set_args(mtry = 4, # size of random subset of variables; default is floor(sqrt(number of total predictors))
           trees = 1000, # Number of trees
           min_n = NULL,
           probability = FALSE, # FALSE: get hard predictions (not needed for regression)
           importance = 'impurity') %>% # we'll come back to this at the end
  set_mode('classification') # change this for regression

# Recipe
data_rec <- recipe(flop ~ No_of_Votes, data = imdb_clean_class) %>%
  step_naomit()

# Workflows
data_wf_mtry5 <- workflow() %>%
  add_model(rf_spec %>% set_args(min_n = 5)) %>%
  add_recipe(data_rec)

## Create workflows for min_n = 10, 25, and 50

data_wf_mtry10 <- workflow() %>%
  add_model(rf_spec %>% set_args(min_n = 10)) %>%
  add_recipe(data_rec)

data_wf_mtry25 <- workflow() %>%
  add_model(rf_spec %>% set_args(min_n = 25)) %>%
  add_recipe(data_rec)

data_wf_mtry50 <- workflow() %>%
  add_model(rf_spec %>% set_args(min_n = 50)) %>%
  add_recipe(data_rec)
```

```{r}
# Fit Models
set.seed(123) # make sure to run this before each fit so that you have the same 1000 trees
data_fit_mtry5 <- fit(data_wf_mtry5, data = imdb_clean_class)

# Fit models for 10, 25, and 50
set.seed(123) 
data_fit_mtry10 <- fit(data_wf_mtry10, data = imdb_clean_class)

set.seed(123)
data_fit_mtry25 <- fit(data_wf_mtry25, data = imdb_clean_class)

set.seed(123) 
data_fit_mtry50 <- fit(data_wf_mtry50, data = imdb_clean_class)
```

```{r}
# Custom Function to get OOB predictions, true observed outcomes and add a user-provided model label
rf_OOB_output <- function(fit_model, model_label, truth){
    tibble(
          .pred_class = fit_model %>% extract_fit_engine() %>% pluck('predictions'), #OOB predictions
          class = truth,
          label = model_label
      )
}
```

@@

data_rf_OOB_output <- bind_rows(
    rf_OOB_output(data_fit_mtry5,5, imdb_clean_class %>% pull(flop)),
    rf_OOB_output(data_fit_mtry10,10, imdb_clean_class %>% pull(flop)),
    rf_OOB_output(data_fit_mtry25,25, imdb_clean_class %>% pull(flop)),
    rf_OOB_output(data_fit_mtry50,50, imdb_clean_class %>% pull(flop))
)

@@

```{r}
rf_OOB_output(data_fit_mtry50,50, imdb_clean_class %>% pull(flop)) %>%
    conf_mat(truth = class, estimate= .pred_class)
```

@@

model_output <-data_fit_mtry50 %>% 
    extract_fit_engine()

model_output %>% 
    vip(num_features = 10) + theme_classic() #based on impurity, 10 meaning the top 10

@@

model_output2 <- data_wf_mtry50 %>% 
  update_model(rf_spec %>% set_args(importance = "permutation")) %>% #based on permutation
  fit(data = imdb_clean_class) %>% 
    extract_fit_engine()

model_output2 %>% 
    vip(num_features = 10) + theme_classic()
